Getting Hitched: Torontonians Prefer to Get Married During Warm Weather

### 2024-01-22 outputs/paper/paper.qmd
additional writing

In order to test this hypothesis, we analyzed the data in the Results section by plotting the number of licenses against the month in which they were obtained and connecting the points with a line so we can see the change in the number of marriage licenses from month to month more clearly. Since there are only 12 data points in each graph, we are also able to lable each observation with the exact number of marriage licenses issued. Then we calculate *****

Our results show that our hypothesis is correct with the number of marriage licenses steadily increasing from Febuary to May, peaking in August, then decreasing between August and November before hitting its lowest in January. This cyclic nature can be observed for both 2022 and 2023

@@

## Cleaned Data
After the data we need has been cleaned, it is saved into 3 .csv files, one for 2022 and another for 2023, both of which have 12 observations one for each month. The third .csv file conains both years and has 24 observations. The only columns that remain in any file are Date and Marriages (@fig-cleaned_toronto_marriages_2022).

@@

cleaned_marriage_data_2022 <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")
gt(head(cleaned_marriage_data_2022))
```

# Results
In order to determine if there is a difference in the amount of marriage licenses issued in Toronto between warm months (April to September) compared to cold months (October to March), we plot the number of marriage licenses for each month separately for 2022, 2023 and both years together. Then the ploints are connected with a path line to make the shape of changes over time easier to see. It is important to note that the lines in the graphs do not represent the actual rate of change between months because there is only one observation for each month. 
## 2022 Marriage Licenses
(@fig-2022_marriages)
```{r}
#| label: fig-2022_marriages
#| fig-cap: Number of marriage licenses issued in Toronto in 2022
#| echo: false
#| message: false
#| eval: true
#| warning: false
marriage_data <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")

marriage_data |>
  ggplot(aes(x = Date, y = Marriages, label = Marriages)) +
  geom_point() +
  geom_path(color = "red") +
  geom_text(hjust=0, vjust=0) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "Month", y = "Marriages", color = "Country")
```
```{r}
#| echo: false
#| message: false
#| include: false
marriage_data_2022 <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")
summary(marriage_data_2022$Marriages)
sum(marriage_data_2022$Marriages)
```

@@

When we examine our graph separated into warm months (April, May, June, July, August, September) and cold months (October, November, December, January, February, March), we can see that there is a 219% increase between the number of marriage licenses in cold and warm months 
$$
Total\ Marriages\ Warm\ Months\ 2022\\
1265+1269+1658+1507+1656+1108=8463
\\
Average=1410.5
\\
Total\ Marriages\ Cold\ Months\ 2022\\
611+476+556+472+536+1220=3871
\\
Average=645.2
\\
$$

@@

marriage_data |>
  ggplot(aes(x = Date, y = Marriages, label = Marriages)) +
  geom_point() +
  geom_path(color = "red") +
  geom_text(hjust=0, vjust=0) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "Month", y = "Marriages", color = "Country")
```

@@

$$
Total\ Marriages\ Warm\ Months\ 2022\\

\\
Average=
\\
Total\ Marriages\ Cold\ Months\ 2022\\

\\
Average=
\\
$$

# Discussion
## Summary Statistics
### 2022
Examining marriage licenses data from 2022, we find a max of 1658, min of 472, and average of 1028, for a combined total of 12334 licenses.

### 2023
In 2023, there was a max of 1649, min of 563, mean of 1058.8, and total of 12706 marriage licenses issued. 
```{r}
#| echo: false
#| message: false
#| include: false
marriage_data_2023 <- read_csv("/cloud/project/outputs/data/cleaned_data_2023.csv")
summary(marriage_data_2023$Marriages)
sum(marriage_data_2023$Marriages)
```

### Both Years
Compared with the 2022 data, the total number of marriage licenses changed by $(12334-12706)/12706 = −0.0292\approx-2.92\%$. This small amount of change indicates that approximately the same number of couples get married in Toronto each year, so any increases in the number of marriages for any given month is unlikely to be attributable to a total increase in marriages.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

>>>>

We formulate our hypothesis to be that there are more marriage licenses obtained in warm months of April to September than cold months of October to March because Torontonians prefer to get married during the warmer half of the year. We analyzed the data in the Results section by plotting the number of licenses against the month in which they were obtained so we can see the change in the number of marriage licenses from month to month. Since there are only 12 data points in each graph, we are also able to label each observation with the exact number of marriage licenses issued. Then we calculated the total amount and the average amount of marriage licenses in the warm months and cold months of each year, and the percent change between them. Our results show that our hypothesis is correct with the number of marriage licenses steadily increasing from February to May, peaking around July, then decreasing between September and November before hitting its lowest in January. This cyclic nature can be observed for both 2022 and 2023.

*** Discussion Section

@@

## Cleaned Data

@@

cleaned_marriage_data_2022 <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")
gt(head(cleaned_marriage_data_2022))
```
After the data we need has been cleaned, it is saved into 3 .csv files, one for 2022 and another for 2023, both of which have 12 observations one for each month. The third .csv file conains both years and has 24 observations. The only columns that remain in any file are Date and Marriages (@fig-cleaned_toronto_marriages_2022).

# Results
In order to determine if there is a difference in the amount of marriage licenses issued in Toronto between warm months (April to September) compared to cold months (October to March), we plot the number of marriage licenses for each month separately for 2022, 2023 and both years together.

## 2022 Marriage Licenses
(@fig-2022_marriages)
```{r}
#| label: fig-2022_marriages
#| fig-cap: Number of marriage licenses issued in Toronto in 2022
#| echo: false
#| message: false
#| eval: true
#| warning: false
marriage_data <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")

marriage_data |>
  ggplot(aes(x = Date, y = Marriages, label = Marriages)) +
  geom_point() +
  geom_text(hjust=0, vjust=0) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "Month", y = "Marriages", color = "Country")
```
```{r}
#| echo: false
#| message: false
#| include: false
marriage_data_2022 <- read_csv("/cloud/project/outputs/data/cleaned_data_2022.csv")
summary(marriage_data_2022$Marriages)
sum(marriage_data_2022$Marriages)
```

@@

When we examine our graph separated into warm months (April, May, June, July, August, September) and cold months (October, November, December, January, February, March), we can see that there is a $(8463-3871)/3871 \approx 117\%$ increase between the number of marriage licenses in cold and warm months 
$$
\begin{aligned}
Total\ Marriages\ Warm\ Months\ &2022\\
1265+1269+1658+1507+1656+1108&=8463
\\
Average&=1410.5
\\
Total\ Marriages\ Cold\ Months\ &2022\\
611+476+556+472+536+1220&=3871
\\
Average&=645.2
\end{aligned}
$$

@@

marriage_data |>
  ggplot(aes(x = Date, y = Marriages, label = Marriages)) +
  geom_point() +
  geom_text(hjust=0, vjust=0) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "Month", y = "Marriages", color = "Country")
```

@@

In 2023 there is a $(7026-4146)/4146\approx68\%$ increase between warm and cold months.

$$
\begin{aligned}
Total\ Marriages\ Warm\ Months\ &2023\\
1086+1507+1507+1628+1649+1156&=7026
\\
Average&=1171
\\
Total\ Marriages\ Cold\ Months\ &2023\\
811+606+647+563+617+929&=4146
\\
Average&=691
\end{aligned}
$$
The results indicate that our hypothesis is correct. There are more marriage licenses issued in warm months (April to September) than cold months (October to March).

## Both Years
```{r}
#| label: fig-marriages
#| fig-cap: Number of marriage licenses issued in Toronto in 2022 and 2023
#| echo: false
#| message: false
#| eval: true
#| warning: false
marriage_data <- read_csv("/cloud/project/outputs/data/cleaned_data_all_years.csv")

marriage_data |>
  ggplot(aes(x = Date, y = Marriages, label = Marriages)) +
  geom_point() +
  geom_path(color = 'red') +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "Month", y = "Marriages", color = "Country")
```
Combining both years, we generate another graph to plot the number of marriage licenses and connect the points with a line so the trend over time is easier to see. It is important to note that the line only connect the points and do not represent the rate of change. This cyclical nature of more marriage licenses obtained in warm months is evident (@fig-marriages)

# Discussion
## Findings
In both 2022 and 2023, there was more marriage licenses issued in warm months than cold ones, 117% more in 2022 and 68% more in 2023. This confirms our hypothesis of more marriages taking place from April to September.

Compared with the 2022 data, the total number of marriage licenses in 2023 only changed by $(12334-12706)/12706 = −0.0292\approx-2.92\%$. This small amount of change indicates that approximately the same number of couples get married in Toronto each year, so any increases in the number of marriages for any given month is unlikely to be attributable to a total increase in marriages.

With these findings we can conclude that April to September is the busy season for marriage license applications and October to March is the less busy season. Since applying for a marriage license comes before the actual wedding ceremony, we can infer that more licenses equate to more weddings. This means that there is more demand for venues, photographers, and catering for weddings from April to September making these services harder to book and potentially more expensive.

Prospective couples should consider the trend of higher volume marriage license applications in the warm months when applying for their marriage license because although many Torontonians prefer getting married in the warm month, the higher volume of marriage license application could cause delays. The higher demand for wedding ceremonies should also be a part of budgeting for their weddings.

## Weaknesses and Next Steps
A weakness in our analysis is the small quantity of data that was used to come to our conclusion. While the pattern of increased marriage licenses was present in both 2022 and 2023, this might not have been the case in prior years. Performing the same analysis on years 2011 to 2021 will confirm our results.

Both years have not only warm monhs where the number of marriage licenses is high and cold months where it is low, but also in between seasons when the number is increasing or decreasing. The analysis could be more nuanced if it had divided the year into 4 instead of 2 by including spring and autumn.

We assumed that obtaining a marriage license comes before conducting the wedding ceremony, but this is not necessarily the case. Some couples can choose to get a license without a ceremony, therefore a higher volume in marriage licenses during warm months does not automatically equate to higher demand for wedding ceremonies. Further analysis of the number of marriage licenses obtained compared with the number of wedding ceremonies that take place should be conducted to confirm this link.
